---
title: "Assignment 3"
output: pdf_document
date: "2026-06-11"
---

```{r}
library(dplyr)
library(tidyr)
library(caret)
library(e1071)

UniversalBank<- read.csv("UniversalBank.csv")

my_data<- UniversalBank

```


```{r}
colnames(my_data)

```

## Part A

```{r}

set.seed(123)

train_index<- createDataPartition(my_data$Personal.Loan, p = 0.6, list = FALSE)

training_data<- my_data[train_index,]
validation_data<- my_data[-train_index,]

```

```{r}

pivot_table<- ftable(CreditCard = training_data$CreditCard, Personal.Loan = training_data$Personal.Loan, Online = training_data$Online)

pivot_table

```
Row variables: CreditCard and Personal.Loan
Column variables: Online
Cell values represent the number of customers in each category combination. 


## Part B

```{r}

57/(475 + 57)

```

P(Personal.Loan = 1 | Online = 1, CreditCard = 1)
Numerator: 57 customers had a credit card, used online banking, and accepted a personal loan.
Denominator: 532 customers had a credit card and used online banking. 
The probability that a customer who owns a bank credit card and is actively using online banking services will acceptr the loan offer is 57/532 = 0.1071, or 10.71%

## Part C

```{r}

online_pivot<- table(Loan = training_data$Personal.Loan, Online = training_data$Online)

online_pivot

```

```{r}

creditcard_pivot<- table(Loan = training_data$Personal.Loan, CreditCard = training_data$CreditCard)

creditcard_pivot

```

## Part D

```{r}

91 / (187 + 91)

179 / (99+179)

278 / 3000

792 / (1930 + 792)

1620 / (1102 + 1620)

2722 / 3000


```
P(CC = 1 | Loan = 1) is 0.3273 or 32.73%
P(Online = 1 | Loan = 1) is 0.6439 or 64.39%
P(Loan = 1) is 0.0927 or 9.27%
P(CC = 1 | Loan = 0) is 0.291 or 29.10%
P(Online = 1 | Loan = 0) is 0.5952 or 59.52%
P(Loan = 0) is 0.9073 or 90.73%

## Part E

```{r}

loan1<- (91 / (187 + 91)) * (179 / (99 + 179)) * (278 / 3000)

loan0<- (792/(1930 + 792)) * (1620 / (1102 + 1620)) * (2722 / 3000)

loan1 / (loan1 + loan0)

```


P(Loan = 1 | CC = 1, Online = 1) is 0.1106 or 11.06%

## Part F

The Naive Bayes probability gave 11.06% while the probability obtained from the pivot table in Part B was 10.71%. Since the Naive Bayes estimate assumes independence between variables, the pivot table value is generally the more accurate estimate. It comes directly from the data without the independence assumption.

## Part G

The entries needed for computing P(Loan = 1 | CC = 1, Online = 1) are:
P(CC = 1 | Loan = 1)
P(Online = 1 | Loan = 1)
P(Loan = 1)
P(CC = 1 | Loan + 0)
P(Online = 1 | Loan = 0)
P(Loan = 0)

These were obtained from the pivot tables created in Part C and were used in the Naive Bayes calculation in Part E.


```{r}

nb_model<- naiveBayes(Personal.Loan ~ CreditCard + Online, data = training_data)

new_customer<- data.frame(CreditCard = 1, Online = 1)

predict(nb_model, new_customer, type = "raw")

```

Running Naive Bayes on the data gave me 0.1157 or 11.57% for P(Loan = 1 | CC = 1, Online = 1), while the probability I obtained in Part E was 0.1106 or 11.06%. The two values are very close and only differ by 0.52 percentage points. This indicates that manual calculation and the Naive Bayes model produce consistent results.













