---
title: "Assignment 1 Dataset from web"
output:
  pdf_document: default
  html_document: default
date: '2026-06-01'
---

# Dataset Source:
# Kaggle - Boost Student Success with College Completion Data
# https://www.kaggle.com/datasets/thedevastator/boost-student-success-with-college-completion-da

## Quantitative Variable

```{r}
summary(data$grade_150_rate)
```

## Qualitative Variable: Gender

```{r}
table(data$gender, data$race)
```


## Variable Transformation
I transformed the quantitative variable `grad_150_rate` into a categorical variable called `completion_level`. Institutions with a graduation rate of 50% or higher were classified as "High Completion" and those below 50% were classified as "Low Completion".

```{r}
data$completion_level <- ifelse(data$grad_150_rate >= 50,
                                "High Completion",
                                "Low Completion")

table(data$completion_level)
```

## Histogram

```{r}
hist(data$grad_150_rate,
     main = "Distribution of 150% Graduation Rates",
     xlab = "Graduation Rate")
```

## Scatterplot

```{r}
plot(data$grad_100_rate,
     data$grad_150_rate,
     main = "100% vs 150% Graduation Rate",
     xlab = "100% Graduation Rate",
     ylab = "150% Graduation Rate")
```

